---
title: "R Notebook"
output: pdf
---
```{r load libraries}
#asdevtools::install_github("tidyverse/googlesheets4")
library(googlesheets4)
library(tidyverse)
# install.packages("devtools")
# the dev version has some workarounds for Google Auth problems
#devtools::install_github("tidyverse/googledrive")
#if(!require(devtools)) install.packages("devtools")
#devtools::install_github("krassowski/complex-upset")
library(googledrive)

library(dplyr)
library(ggalluvial)
library(xtable)
library(naniar)
library(sqldf)
library(VennDiagram)
library(magrittr)
library(here) #https://www.tidyverse.org/blog/2017/12/workflow-vs-script/
library(ComplexUpset)

```

```{r load_data}
#coding_sheet = drive_get("2019 recoding")
load_emse_from_google = function() {
        data.emse_all = read_sheet(coding_sheet, sheet="emse2017", range="B2:AF85", col_names=TRUE)
        data.emse = data.emse_all %>% select(c(2,3,5,24,25,26,27,28,29,30,31))
        colnames(data.emse) = c("ID","title","authors", "strategy1","strategy2","quadrant","purpose","human","system","researcher","practitioners")
        cols_to_factor = c("strategy1","strategy2","quadrant","purpose","human","system", "researcher","practitioners")
        data.emse %<>% mutate_at(cols_to_factor, funs(factor(.)))
        data.emse
}

load_icse2_from_google = function() {
        icse2.data = read_sheet(coding_sheet, sheet="icse2017", range="A2:AE70", col_names=TRUE)
        data.icse2 = icse2.data %>% select(c(3,4,5,24,25,26,27,28,29,30,31))
        colnames(data.icse2) = c("ID","title","authors", "strategy1","strategy2","quadrant","purpose","human","system","researcher","practitioners")
        cols_to_factor = c("strategy1","strategy2","quadrant","purpose","human","system", "researcher","practitioners")
        data.icse2 %<>% mutate_at(cols_to_factor, funs(factor(.)))
        data.icse2
}

load_from_disk = function(filename) {
  data = read.csv(here("data",filename)) #here() figures out relative paths
} 

# EMSE 2017
emse = load_from_disk("emse2017_coded_data_0123.csv")
#emse = load_emse_from_google() # only if Google coding has changed
emse_total = nrow(emse)
# write.csv(emse,here("data","emse2017_coded_data_0123.csv"))
## ICSE 2017
# icse = load_icse2_from_google()
icse = load_from_disk('icse2017-coded_data0120.csv')
icse_total = nrow(icse)
```

```{r Create Quadrants}
emse_df = subset(emse, select= -c(ID,title,authors,quadrant))#removes title/author/id etc.
icse_df = subset(icse, select= -c(ID,title,authors,quadrant))

emse_df = emse_df %>% mutate(venue = 'EMSE')
icse_df = icse_df %>% mutate(venue = 'ICSE')
both_df = bind_rows(icse_df,emse_df) #merge into one df

both_df$human = as.logical(both_df$human)
both_df$researcher = as.logical(both_df$researcher)
both_df$system = as.logical(both_df$system)

# create the new beneficiary column as boolean of the 3 original cols
both_df = both_df %>% mutate(beneficiary = case_when( 
  human & !researcher & !system ~ 'Human',
  !human & researcher & !system ~ 'Researcher',
  !human & !researcher & system ~ 'System',
  human & system & !researcher ~ 'Human_System',
  human & researcher & !system ~ 'Human_Researcher',
  !human & researcher & system ~ 'System_Researcher',
  human & researcher & system ~ 'All'))

both_df$beneficiary = as.factor(both_df$beneficiary)

human_strats = c("JudgmentStudy","SampleSurvey","FieldStudy","FieldExperiment","LaboratoryExperiment","ExperimentalSimulation")
drop_strats = c("FormalTheory", "Meta") # remove these entries to re-instate meta
# calculate mixed studies. This is the 'quadrant' column in the spreadsheet, but this gives us 
# flexibility for the plotting.
both_df = both_df %>% mutate(quadrant = case_when(
  strategy1 == "ComputationalStudy" & (is.na(strategy2) | strategy2 == "ComputationalStudy" | strategy2 %in% drop_strats) ~ "Data",
  strategy2 == "ComputationalStudy" & strategy1 %in% drop_strats ~ "Data",

  strategy1 == "ComputationalStudy" & strategy2 %in% human_strats ~ "HumanData",
  strategy2 == "ComputationalStudy" & strategy1 %in% human_strats ~ "HumanData",

  strategy1 == "JudgmentStudy" & (is.na(strategy2) |  strategy2 %in% drop_strats) ~ "Respondent",
  strategy1 == "SampleSurvey" &  (is.na(strategy2) | strategy2 %in% drop_strats) ~ "Respondent",
  strategy1 == "FieldExperiment" & (is.na(strategy2) | strategy2 %in% drop_strats) ~ "Field",
  strategy1 == "FieldStudy" & (is.na(strategy2) | strategy2 %in% drop_strats) ~ "Field",
  strategy1 == "ExperimentalSimulation" & (is.na(strategy2) | strategy2 %in% drop_strats) ~ "Lab",
  strategy1 == "LaboratoryExperiment" & (is.na(strategy2) | strategy2 %in% drop_strats) ~ "Lab",

  strategy2 == "JudgmentStudy" & strategy1 %in% drop_strats ~ "Respondent",
  strategy2 == "SampleSurvey" & strategy1 %in% drop_strats ~ "Respondent",
  strategy2 == "FieldExperiment" &  strategy1 %in% drop_strats ~ "Field",
  strategy2 == "FieldStudy" & strategy1 %in% drop_strats ~ "Field",
  strategy2 == "ExperimentalSimulation" &  strategy1 %in% drop_strats ~ "Lab",
  strategy2 == "LaboratoryExperiment" &  strategy1 %in% drop_strats ~ "Lab",
  
  strategy2 %in% human_strats & strategy1 %in% human_strats ~ "HumanHuman",
  
  # (strategy1 == "Meta" | strategy2 == "Meta") & is.na(strategy2) ~ "Meta",
  strategy1 == "FormalTheory" & is.na(strategy2) ~ "Theory",
  strategy1 == "Meta" & is.na(strategy2) ~ "Meta",

  # for the paper, leave out FormalTheory and Meta studies entirely
  # strategy2 == "FormalTheory" ~ NA,
  # strategy2 == "Meta" ~ NA,
))
```
"all" mixed should be 53 (strat1 & strat2 have entries)
nrow(both_df %>% filter(!is.na(strategy2)))
mixed with Meta = 2 with Strat2 = meta and 1 with Strat1 = Meta with something else
nrow(both_df %>% filter(!is.na(strategy2) & strategy1 == "Meta"))
mixed, formalTheory = 9 where FormalTHeory is strat2, 3 where FormalTheory = Strat1 with something else Strat2
so 12 mixed = FormalTHeory, 3 mixed are Meta
and quadrant=mixed = 41, after applying mutate above

```{r create alluvia}

both_df = subset(both_df, select=-c(human, system,researcher,strategy1,strategy2)) # don't need these boolean cols anymore

alluvia = both_df %>% 
  group_by(quadrant,purpose,beneficiary,venue) %>% 
  summarize(count=n()) %>%
  mutate(proportion = case_when(
     venue == 'EMSE' ~ count/emse_total,
     venue == 'ICSE' ~ count/icse_total
  ))
```

```{r RQ1 - beneficiary}
#TODO change table so it is All/ICSE/EMSE style
ben_df = alluvia %>% group_by(beneficiary,venue) %>% summarize(count = sum(count))
print(xtable(ben_df),booktabs=TRUE,include.rownames = FALSE)

ggplot(ben_df, aes(x=beneficiary, y=count)) +
  geom_col(aes(fill=venue) , position="dodge2") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank()) +
  scale_fill_manual("venue", values = c("EMSE" = "orange", "ICSE" = "turquoise4"))+
  geom_text(aes(label=count), position=position_dodge2(width=0.9), vjust=-0.15)

ggsave(here("fig","both_benefits_bar.pdf"), width=6, height=4)

```

```{r Strategies for fig2}
quad_df = alluvia %>% group_by(quadrant,venue) %>% summarize(count = sum(count))
print(xtable(quad_df),booktabs=TRUE,include.rownames = FALSE)
#TODO extend without mixed
```



```{r RQ2 find totals for Purpose/Strategy}
purpose_df = alluvia %>% 
  group_by(purpose,venue) %>% 
  summarize(count = sum(count)) %>% 
  mutate(proportion = case_when(
     venue == 'EMSE' ~ count/emse_total,
     venue == 'ICSE' ~ count/icse_total
  ))
print(xtable(purpose_df),booktabs=TRUE,include.rownames = FALSE)

```

```{r RQ4}
# plot flow map
#https://cran.r-project.org/web/packages/ggalluvial/vignettes/ggalluvial.html
simplified = both_df %>% 
    filter(beneficiary != 'Researcher') %>% # drop
    filter(  quadrant!='Meta' & quadrant != 'Theory') %>% 
    mutate(beneficiary = case_when(
      beneficiary == "Human_Researcher" ~ 'Human',
      beneficiary == "System_Researcher" ~ 'System',
      beneficiary == "Human" ~ 'Human',
      beneficiary == "System" ~ 'System',
      beneficiary == "Human_System" ~ 'Human_System',
      beneficiary == "Human_Researcher" ~ 'Human',
      beneficiary == "All" ~ 'Human_System'
    )) # this one drops researcher and re-bins those

simplified_sum = simplified %>% 
  group_by(quadrant,purpose,beneficiary,venue) %>% 
  summarize(count=n()) %>%
  mutate(proportion = case_when(
     venue == 'EMSE' ~ count/emse_total,
     venue == 'ICSE' ~ count/icse_total
  ))

simplified_sum$quadrant = factor(simplified_sum$quadrant,levels=c("Respondent","Field","Lab","HumanHuman","HumanData","Data"))
icse_simplified = simplified_sum %>% filter(venue == 'ICSE')
emse_simplified = simplified_sum %>% filter(venue == 'EMSE')
```

``` {r plot alluvial}
plot_alluvial = function(filename) {
  ggplot(filename,  
       aes(y=count, 
           axis1=beneficiary, 
           axis2=quadrant, 
           axis3=purpose)) +
  geom_alluvium(aes(fill=quadrant),width=1/12) +
  geom_stratum(width=1/12, fill="grey", color ="black") +
  geom_label(stat="stratum",infer.label=TRUE,size=2) +
  scale_x_discrete(labels = c("Beneficiary\n(Who)","Quadrant\n(How)","Purpose\n(What)"),limits=c("Beneficiary","Quadrant","Purpose"))+
  scale_fill_brewer(palette="Spectral", direction=-1)+
  # scale_fill_viridis(discrete=TRUE) +
   ylab("Number of Papers") +
  theme_bw()
  # ggtitle("Research Papers - What, How, Who")
}
plot_alluvial(icse_simplified)
ggsave(here("fig","icse_alluvial.pdf"), width=6, height=4)
plot_alluvial(emse_simplified)
ggsave(here("fig","emse_alluvial.pdf"), width=6, height=4)
```


``` {r Practitioner}
practitioner_df = both_df %>% group_by(practitioners, venue) %>% summarize(count = n()) %>% mutate(proportion = case_when(
     venue == 'EMSE' ~ count/emse_total,
     venue == 'ICSE' ~ count/icse_total
))
print(xtable(practitioner_df),booktabs=TRUE,include.rownames = FALSE)

```

```{r Triangulation and RQ3 Strategy}
strats_emse = emse[c("strategy1","strategy2")]
strats_icse = icse[c("strategy1","strategy2")]
strats_emse = strats_emse %>% mutate(venue = 'EMSE')
strats_icse = strats_icse %>% mutate(venue = 'ICSE')
both_strats = bind_rows(strats_icse,strats_emse) #merge into one df
strats_summary = both_strats %>%
  group_by(strategy1,strategy2, venue) %>%
  summarize(count = n())
comp_study = strats_summary %>% 
  filter(strategy1 == "ComputationalStudy" | strategy2 == "ComputationalStudy" ) %>%
  group_by(venue) %>%
  summarize(total = sum(count))

# double-counts papers with two strategies!
all_strats = both_strats %>% pivot_longer(cols=starts_with("stra"), values_drop_na=TRUE,names_to="strategy")
vis = all_strats %>%
  group_by(venue,value) %>%
  summarize(count=n())
print(xtable(vis),booktabs=TRUE,include.rownames = FALSE)
```

Now print the totals for quadrants, double counting papers with two strategies
``` {r quadrant totals}
new_all_strats = map_strategy_quadrant(all_strats, 'quadrant','value')
new_all_strats = new_all_strats %>% group_by(quadrant,venue) %>% summarize(count = n())
print(xtable(new_all_strats),booktabs=TRUE,include.rownames = FALSE)

```

``` {r triangulation by quadrant}
triangulate_df = bind_rows(icse_df,emse_df) #merge into one df
triangulate_df = map_strategy_quadrant(triangulate_df, 'quadrant1', 'strategy1')
triangulate_df = map_strategy_quadrant(triangulate_df, 'quadrant2', 'strategy2')
# 
# ))  %>% mutate(quadrant2 = case_when(
#   strategy2 == "ComputationalStudy" ~ "Data",
#   strategy2 == "JudgmentStudy" | strategy2 == "SampleSurvey" ~ "Respondent",
#   strategy2 == "FieldExperiment"  | strategy2 == "FieldStudy" ~ "Field",
#   strategy2 == "ExperimentalSimulation" | strategy2 == "LaboratoryExperiment" ~ "Lab",
#   strategy2 == "FormalTheory" ~ "Theory",
#   strategy2 == "Meta" ~ "Meta"
triangulate_df = triangulate_df %>% 
  group_by(venue,quadrant1,quadrant2) %>% 
  summarize(count = n())

print(xtable(triangulate_df),booktabs=TRUE,include.rownames = FALSE)
```

``` {r utility functions}
map_strategy_quadrant = function(df, new_col, compare_col) {
  new_df = df %>% mutate(!!new_col := case_when(
  get(compare_col) == "ComputationalStudy" ~ "Data",
  get(compare_col)  == "JudgmentStudy" | get(compare_col)  == "SampleSurvey" ~ "Respondent",
  get(compare_col)  == "FieldExperiment"  | get(compare_col)  == "FieldStudy" ~ "Field",
  get(compare_col)  == "ExperimentalSimulation" | get(compare_col)  == "LaboratoryExperiment" ~ "Lab",
  get(compare_col)  == "FormalTheory" ~ "Theory",
  get(compare_col)  == "Meta" ~ "Meta"
))
}
```



``` {r plotRQ3}
all_strats$value = recode(all_strats$value, "ComputationalStudy"="Data")
all_strats$value = factor(all_strats$value, levels=c("SampleSurvey","JudgmentStudy","LaboratoryExperiment","ExperimentalSimulation",
                                                        "FieldExperiment","FieldStudy","Meta","FormalTheory","Data"))

ggplot(all_strats,aes(value)) +
  geom_bar(aes(fill=venue),position = "dodge2") + 
  theme_bw() + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_fill_manual("venue", values = c("EMSE" = "orange", "ICSE" = "turquoise4"))+
  geom_text(stat="count",aes(fill=venue,label=..count..), position=position_dodge2(width=0.9), vjust=-0.15)


ggsave(here("fig","both_strategies_bar.pdf"), width=6, height=4)

```

How many ICSE PC/PB members also sit on the EMSE editorial board?
``` {r overlap of PC/EB}
emse_ed = read.csv(here("data","/emse2020-eds.csv"),header=FALSE)
icse_pc = read.csv(here("data","/icse2020-pc.csv"), header=FALSE)
emse_ed$in_emse = TRUE
icse_pc$in_icse = TRUE
res = merge(emse_ed,icse_pc,all=TRUE)
match = res %>% filter(in_emse & in_icse)

```


```{r plot venn beneficiaries}
# TODO change the data to refer to ben_df
# obsolete - see venn.ipynb using matplotlib venn package
# needs 3 sets of  (Abc, aBc, ABc, abC, AbC, aBC, ABC)
# All A Human Human_Researcher  Human_System  B Researcher  C System  System_Researcher
venn_data = both_df %>% select(beneficiary,venue)
# set_icse = venn_data %>% filter(venue=="ICSE")
# set_emse = venn_data %>% filter(venue=="EMSE")
# set_icse = set_icse %>% select(beneficiary)
# set_emse = set_emse %>% select(beneficiary)
# write.csv(set_emse, here("data","emse_venn.csv"),row.names=FALSE)
# write.csv(set_icse, here("data","icse_venn.csv"),row.names=FALSE)

#venn_data = {}
# venn_data["ABC"] = ben_df[1,3] + ben_df[2,3]
# venn_data["Abc"] = ben_df[3,3] + ben_df[4,3]
# venn_data["aBc"] = ben_df[9,3] + ben_df[10,3]
# venn_data["abC"] = ben_df[11,3] + ben_df[12,3]
# venn_data["ABc"] = ben_df[5,3] + ben_df[6,3]
# venn_data["aBC"] = ben_df[13,3] + ben_df[14,3]
# venn_data["AbC"] = ben_df[7,3] + ben_df[8,3]

# venn.plot = draw.triple.venn(area1 = venn_data$Abc, area2 = venn_data$aBc, area3 = venn_data$abC,
#                              n12 = venn_data$ABc, n23 = venn_data$aBC, n13 = venn_data$AbC, 
#                              n123 = venn_data$ABC, # intersection of all three 
#                            category = c("Human", "System", "Researcher"), 
#                              lty = "blank",fill = c("skyblue", "pink1", "mediumorchid"))
# tiff(filename = "Triple_Venn_diagram.tiff", compression = "lzw");grid.draw(venn.plot);dev.off();
```

``` {r Upset Applied to data}

venn_data = both_df %>% select(beneficiary,venue)
cats = c("Human","Researcher","System")
names(cats) = cats
venn_data$beneficiary = venn_data$beneficiary %>% recode(All = 'Human_System_Researcher')
ben_mat <- map_dfc(venn_data$beneficiary, str_detect, cats) %>%
    data.frame() %>%
    t() %>% 
    as_tibble()
ben_mat = ben_mat %>% add_column(Venue = venn_data$venue)
# ben_mat = mutate(ben_mat, Venue = case_when(Venue == "ICSE" ~ TRUE, Venue == "EMSE" ~ FALSE))

colnames(ben_mat)  <- c(cats,"Venue")

upset(data = ben_mat, 
      intersect = cats,
      set_sizes = FALSE,
      annotations = list(
        'Venue'=list(
            aes=aes(x=intersection, fill=Venue),
            geom=list(
                geom_bar(stat='count', position='fill'),
                scale_y_continuous(labels=scales::percent_format()),
                scale_fill_manual(values=c(
                    'ICSE'='#E41A1C', 'EMSE'='#377EB8'
                ))
            )
        )
      ),
      # set_sizes=upset_set_size(
      #       width=0.4
      # ),
      # themes=upset_modify_themes(
      #   list(
      #       'overall_sizes'=theme(axis.text.x=element_text(angle=90))
      #   )
      # ),
      name="Beneficiary Groupings by Frequency. Total is 151 papers.",
      min_size = 0,
      width_ratio = 0.125) + 
      labs(title = "Co-occurrence of Beneficiaries",
         caption = "Idea: @kjhealy/ComplexUpset")
# upset_query(set=Venue_ICSE, color='red', fill='red')
ggsave(here("fig","upset.pdf"), width=6, height=4)

```